---
title: "R Data Structures"
author: "Bryan Adams"
date: "November 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = F, messages = F)

library(Matrix)
library(tidyverse)
library(purrr)

```

# Types of Data Structures


1. Homogeneous	
* 1d	Atomic vector	
* 2d	Matrix	
* nd	Array	

2. Heterogeneous
*	List
*	Data frame

# Homogeneous

## Vectors

```{r}
vector = c(1.1,1.2)
str(vector)
```

```{r}
vector = c("a","b")
str(vector)
```

### Factors

#### What is a factor and why we should use it.

```{r}
sex = c("M","F","M","F")
str(sex)

```

I just made a character vector with a list of sexes

Now I will make it a "factor"

```{r}
sex_factor = factor(sex)
str(sex_factor)
```

The size changes too.

```{r}
object.size(sex)
object.size(sex_factor)
```


```{r}
levels(sex_factor)
```

Maybe I want to rename them

```{r}
levels(sex_factor) = c("Female","Male")
sex_factor
```

I can also break them


```{r}
sex_factor = c(sex_factor, "Male")
str(sex_factor)
```

## Matrix vs Sparse Matrix

```{r}
library(Matrix)
```

```{r}
m1 = matrix(0, nrow = 10000, ncol = 10000)
m2 = Matrix(0, nrow = 10000, ncol = 10000, sparse = TRUE)
```

```{r}
str(m1)
object.size(m1)
```

```{r}
str(m2)
object.size(m2)
```

The full matrix representation does not change in size because all of the zeros are being represented explicitly, while the sparse matrix is conserving that space by representing only the non-zero entries. 

```{r}
m1 = matrix(0, nrow = 100000, ncol = 100000)
m2 = Matrix(0, nrow = 100000, ncol = 100000, sparse = TRUE)
```

## Array

In arrays, data is stored in the form of matrices, rows, and columns. We can use the matrix level, row index, and column index to access the matrix elements.

```{r}
my_array = array(1:12, c(2, 2, 3))

my_array

dim(my_array)

str(my_array)
```

Here the array has the following diminsions:  2 - rows, 2 - columns, 3 - matrices.

```{r}
column_names <- c("COL1","COL2")
row_names <- c("ROW1","ROW2")
matrix_names <- c("Matrix1","Matrix2", "Matrix3")

my_array = array(1:12, dim = c(2,2,3), dimnames = list(row_names, column_names, matrix_names))

my_array
```

# Heterogenous

## Lists

A list in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, and even other lists

### Making a tibble

```{r}
tib_1 = tibble(
  x = 1:5,
  y = 1:5,
  z = x+y
)


tib_2 = tibble(
  x = 10:15,
  y = 10:15,
  z = x+y
)

str(tib_2)
```

```{r}
my_tibs = list(tib_1,tib_2)

str(my_tibs)

my_tibs%>%
  map(~summary(.x))

```

### Lists can have anything

```{r}
my_list = list(my_tibs, sex)
str(my_list)
```

# Why?

```{r}
Sys.time()
Sys.Date()

c(Sys.time(),
Sys.Date())

```